---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sunflower: Assessing and Categorizing Production Errors in Spanish

<!-- badges: start -->
<!-- badges: end -->

The goal of *sunflower* is to handle multiple response data, compute formal metrics, and classify production errors, whether in speech or spelling transcriptions. The outputs of this package make it easy to run statistical analyses in [R](https://www.r-project.org/).

## Installation

You can install the current version of *sunflower* from [GitHub](https://github.com/) using the following code lines.
Also, it is recommended to install the [tidyverse package](https://www.tidyverse.org/) to allow the work with pipes, 
and the [word2vec CRAN package](https://cran.r-project.org/web/packages/word2vec/readme/README.html) for computations involving word2vec models required to classify errors (as described in the last section of this markdown).

```r
install.packages("devtools")
devtools::install_github("ismaelgutier/sunflower")
install.packages("tidyverse")
```
Once these packages have been installed, it is only required to load then in R (recommended working with [RStudio](https://posit.co/download/rstudio-desktop/) to have an optimal IDE).

```r
require("sunflower")
require("tidyverse")
```

## How to use

```{r include=FALSE}
require("sunflower")
require("tidyverse")
require("htmltools")
```

### Compute formal similarity metrics

```{r}
df_to_formal_metrics = sunflower::IGC_long_phon %>% select(-c(modality, task_modality,task_type, test, task))


formal_metrics_computed = df_to_formal_metrics %>% get_formal_metrics(item_col = "item_phon",
                                             response_col = "response_phon",
                                             attempt_col = "Attempt",
                                             group_cols = c("ID", "item_ID"))

formal_metrics_computed %>% head(8) %>% knitr::kable()

```
`Note`: Move the dataframe to the right to see all the columns and metrics.


### Positional accuracy

```{r message=FALSE}
positions_accuracy = formal_metrics_computed %>% 
  position_scores(match_col = "itemL_adj_strict_match_pos", last_ID_col = "targetL")

positions_accuracy %>% head(8) %>% knitr::kable()
```
```{r include=FALSE}
# Convertir targetL a character para evitar problemas al combinar dataframes
positions_accuracy <- positions_accuracy %>%
  mutate(targetL = as.character(targetL))

# Duplicar y modificar el dataframe para crear 'positions_general'
positions_general <- positions_accuracy %>%
  mutate(targetL = "General")

# Combinar ambos dataframes
positions <- bind_rows(positions_accuracy, positions_general)

# Especificar manualmente los niveles en el orden deseado
desired_levels <- c("3", "4", "5", "6", "7", "8", "9", "10", "11", "12",
                    "13", "14", "15", "17", "21", "22", "24", "48", "General")

# Convertir correct_pos a numérico y ordenar targetL como factor según desired_levels
positions <- positions %>%
  mutate(correct_pos = as.numeric(correct_pos),
         targetL = factor(targetL, levels = desired_levels)) %>%
  arrange(correct_pos, targetL)

# Definir un conjunto de linetypes que se pueda repetir
custom_linetypes <- rep(c("solid", "dashed", "dotted", "longdash", "dotdash"),
                        length.out = nlevels(positions$targetL))

# Calcular la precisión y contar el número de observaciones por grupo
plot_positions <- positions %>%
  group_by(Position, targetL) %>%
  summarize(acc = mean(correct_pos, na.rm = TRUE),
            n = n()) %>%
  ggplot(aes(x = as.numeric(Position), y = acc, group = targetL,
             fill = targetL, color = targetL, lty = targetL)) +
  geom_line(size = 0.70, alpha = 0.6) +
  geom_point(aes(size = n), shape = 21, color = "black", alpha = 0.6) +
  scale_linetype_manual(values = custom_linetypes) +
  theme(panel.border = element_rect(colour = "black", fill = NA),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank()) +
  ylab("Proportion (%) of correct phonemes") +
  xlab("Phoneme position") +
  guides(fill = guide_legend(title = "Word Length"),
         lty = guide_legend(title = "Word Length"),
         color = guide_legend(title = "Word Length"),
         size = guide_legend(title = "Datapoints")) + 
  theme_gray() +
  theme(legend.position = "right",
        legend.key.size = unit(0.6, "lines"))

```

`Note`: A plot depicting the positions' accuracy of 14,418 datapoints.

```{r plot_positions, out.width = "75%", fig.align="center", echo=FALSE}
plot_positions
```

### Classify productions 

```{r include=FALSE}
df_to_classify = sunflower::IGC_long %>% select(-c(modality, task_modality,task_type, test, task))

wordlist <- read_csv(file = file.choose(), 
                     col_names = FALSE, 
                     show_col_types = FALSE, 
                     locale = locale(asciify = TRUE)) %>% as.vector()

m_w2v = word2vec::read.word2vec(file = file.choose(), normalize = TRUE)
```

```{r}
errors_classified = df_to_classify %>% 
  get_formal_similarity_indexes(target_col = "item", response_col = "Response", 
                            item_type = "task_type", source1 = wordlist) %>%
  get_cosine_similarity_df(target_col = "item", response_col = "Response", model = m_w2v) %>%
  classification(access_col = "accessed", RA_col = "RA")

errors_classified %>% head(8) %>% knitr::kable()
```

`Note 1`: Move the dataframe to the right to see all the columns and errors.

`Note 2`: The `source1 = wordlist` is a txt file contained in the dependency-bundle zip that can be found at our supplementary [OSF repository](https://osf.io/akuxv/). The `model = m_w2v` is a word2vec binary file located also in that zip (see the markdown in the vignettes for further info).

----


Any suggestions, comments, or questions about the functionality of the package are warmly welcomed. If you are interested in contributing to the project, such as by expanding it to other languages, please feel free to contact us.

🌻
