---
title: "Data Analysis Workflow using the Sunflower Package"
author: "Ismael Gutiérrez-Cordero"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

In this vignette, we present a practical example of using the sunflower package to work with datasets that include a column of responses containing multiple answers. We demonstrate how to convert the dataset into a long format to obtain formal similarity metrics. Additionally, we illustrate how to perform error classification based on classical criteria found in the literature (e.g., [Dell et al., 1997](https://doi.org/10.1037/0033-295x.104.4.801); [Gold & Kertesz, 2001](https://doi.org/10.1006/brln.2000.2441); see also, [García-Orza et al., 2020](https://doi.org/10.1016/j.cortex.2020.03.020)).

# Environment Setup

```{r}
# Clear the workspace and unload all packages
#rm(list = ls())
#invisible(lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""),
#                 detach, character.only = TRUE, unload = TRUE))

# Install and load `devtools` package
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
library(devtools)

# Install RTools on Windows (if applicable)
# Visit: https://cran.rstudio.com/bin/windows/Rtools/ for installation.
# Not needed on macOS or Linux (I am not an user, so I am guessing)
```

# Install and Load Required Packages

```{r}
# Install additional packages if needed
possible_dependencies <- c("tidyverse", "htmlTable", "knitr")
for (pkg in possible_dependencies) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}

# Install the `xfun` package (if necessary)
if (!requireNamespace("xfun", quietly = TRUE)) {
  install.packages("xfun", type = "source")
}

# Install the `sunflower` package from GitHub
if (!requireNamespace("sunflower", quietly = TRUE)) {
  devtools::install_github("ismaelgutier/sunflower")
}

# Load required libraries
library(sunflower)
```

# Step 1
## Load and Wrangle Data

```{r}
# Load dataset
dataframe0 <- sunflower::IGC_sample

# Separate responses
dataframe1 <- dataframe0 %>% 
  sunflower::separate_responses(col_name = "response",
                                separate_with = ", ")

# Extract attempts and clean blank spaces
dataframe2 <- dataframe1 %>% 
  sunflower::get_attempts(first_production = attempt_1, 
                          drop_blank_spaces = TRUE)
```

# Step 2
## Formal Similarity Analysis

```{r}
# Calculate formal similarity
dataframe3 <- dataframe2 %>%
    sunflower::get_formal_similarity(item_col = "item",
                                     response_col = "response",
                                     attempt_col = "attempt",
                                     group_cols = c("task_item_ID"))
```

# Step 2.1
## Positional Accuracy

```{r}
# Calculate positional accuracy
dataframe3.1 <- dataframe3 %>% 
    sunflower::positional_accuracy(item_col = "item", 
                                   response_col = "response",
                                   match_col = "adj_strict_match_pos")
```

# Step 3
## Lexicality Check

```{r}
# Check lexicality
dataframe4 <- dataframe3 %>%
    sunflower::check_lexicality(item_col = "item",
                                response_col = "response",
                                criterion = "dictionary")
```

## Semantic Similarity Analysis

```{r}
# Load a pre-trained word2vec model
model <- word2vec::read.word2vec(file = file.choose(), normalize = FALSE)

# Calculate semantic similarity
dataframe5 <- dataframe4 %>%
    sunflower::get_semantic_similarity(item_col = "item",
                                       response_col = "response",
                                       model = model)
```

## Error Classification

### Classify Errors Considering Retrieval Attempts (RAs)

```{r}
dataframe6a <- dataframe5 %>%
  dplyr::select(-correct) %>% # remove old correct_column (the user might also rename if want to keep it)
  dplyr::mutate(accessed = ifelse(item == response, 1, 0)) %>%
  sunflower::classify_errors(access_col = "accessed", 
                  RA_col = "RA",
                  response_col = "response", 
                  item_col = "item",
                  also_classify_RAs = TRUE,
                  cosine_limit_value = 0.46)
knitr::kable(dataframe6a) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  kableExtra::scroll_box(width = "120%", height = "500px")
```

### Classify Errors Without Considering RAs

```{r}
dataframe6b <- dataframe5 %>%
  dplyr::select(-correct) %>% # remove old correct_column (the user might also rename if want to keep it)
  dplyr::mutate(accessed = ifelse(item == response, 1, 0)) %>%
  sunflower::classify_errors_regular(access_col = "accessed", 
                          response_col = "response", 
                          item_col = "item",
                          cosine_limit_value = 0.46)
knitr::kable(dataframe6b) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  kableExtra::scroll_box(width = "120%", height = "500px")

```

# Conclusion

This R Markdown document provides a complete workflow for analyzing data using the sunflower package, incorporating data wrangling, similarity metrics, and error classification.

# R Session Info.
```{r}
sessionInfo()
```
