---
title: "Full working procedure using `sunflower`"
author: "Ismael Gutiérrez-Cordero"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

In this vignette, we present a practical example of using the sunflower package to work with datasets that include a column of responses containing multiple answers. We demonstrate how to convert the dataset into a long format to obtain formal similarity metrics. Additionally, we illustrate how to perform error classification based on classical criteria found in the literature (e.g., [Dell et al., 1997](https://doi.org/10.1037/0033-295x.104.4.801); [Gold & Kertesz, 2001](https://doi.org/10.1006/brln.2000.2441); see also, [García-Orza et al., 2020](https://doi.org/10.1016/j.cortex.2020.03.020)).

# 0. Dependencies

```{r dependencies}
require(sunflower) # to work
require(tidyverse) # to work along sunflower 
require(knitr) # to work in Rmarkdown
require(kableExtra) # to work in Rmarkdown
require(rmarkdown) # to work in Rmarkdown
require(ggplot2) #to plot
```

# 1. Managing multiple responses

Load the data `IGC` allocated in the package, select some columns to keep and visualize it.

```{r, view data}

IGC = sunflower::IGC

IGC = IGC %>% dplyr::select(task_ID, task_type, ID, task_item_ID, item, final_response, correct) %>% dplyr::filter(task_ID == "SnodgrassVanderwart_nam")

rmarkdown::paged_table(IGC %>% dplyr::select(c(ID, task_item_ID, item, final_response)), options = list(rows.print = 8, align = "ccc"))

  # Convertir el dataframe en un objeto tipo tabla usando tableGrob
  table_plot <- gridExtra::tableGrob(IGC %>% dplyr::select(c(ID, task_item_ID, item, final_response)) %>%
    dplyr::slice_head(n = 7))
  
  # Crear un gráfico vacío para agregar la tabla, eliminando los márgenes
  plot <- ggplot() +
    theme_void() +  # Eliminar todos los elementos gráficos
    annotation_custom(table_plot)
  
  # Mostrar el gráfico
  plot
  
  ggsave("f1_initial.png", plot, width = 8, height = 4.15, dpi = 600)

```

Separate the data using the `separate_responses()` function and them rearrange as long format using the `join_responses()` to work in the following step. 

```{r, long data}

IGC_step1 = IGC %>% separate_responses(col_name = "final_response",
                            separate_with = ", ") %>% 
                        get_attempts(first_production = Attempt_1, drop_blank_spaces = T)

IGC_step1_skinnydf = IGC_step1 %>% dplyr::select(-c(task_ID, task_type))

rmarkdown::paged_table(IGC_step1_skinnydf, options = list(rows.print = 25, align = "ccc"))
  
  # Convertir el dataframe en un objeto tipo tabla usando tableGrob
    table_plot2 <- gridExtra::tableGrob(IGC_step1_skinnydf %>% dplyr::select(c(ID, item_ID = task_item_ID, item, attempt = Attempt, response = Response)) %>%
      dplyr::slice_head(n = 7))
    
    # Crear un gráfico vacío para agregar la tabla, eliminando los márgenes
    plot2 <- ggplot2::ggplot() +
      ggplot2::theme_void() +  # Eliminar todos los elementos gráficos
      ggplot2::annotation_custom(table_plot2)
    
    # Mostrar el gráfico
    plot2
    
    ggsave("f1_second.png", plot2, width = 8, height = 4.15, dpi = 600)
    
```

# 2. Formal Analysis

Compute the similarity metrics using the `get_formal_similarity()` function.

```{r, formal metrics}

IGC_step2 = IGC_step1 %>% get_formal_similarity(item_col = "item",
                                               response_col = "Response",
                                               attempt_col = "Attempt",
                                               group_cols = c("ID", "task_item_ID"))

IGC_step2_skinnydf = IGC_step2 %>% dplyr::select(-c(task_ID, task_type))

rmarkdown::paged_table(IGC_step2_skinnydf, options = list(rows.print = 25, align = "c"))


  # Convertir el dataframe en un objeto tipo tabla usando tableGrob
      table_plot3 <- gridExtra::tableGrob(
        IGC_step2_skinnydf %>% dplyr::select(-c(responseL, targetL)) %>%
            dplyr::mutate(across(where(is.numeric), ~ round(., 3))) %>% 
            dplyr::select(c(task_item_ID, item, Response, RA, Attempt, shared1char:DLd, JWd, pcc, CdA_diff = approach_diff)) %>%
          dplyr::rename(item_ID = task_item_ID, diff_chars = diff_char_num, attempt = Attempt, response = Response) %>%
          dplyr::slice_head(n = 7) # Seleccionar las primeras 7 filas
      )

  # Crear un gráfico vacío para agregar la tabla, eliminando los márgenes
    plot3 <- ggplot() +
      theme_void() +  # Eliminar todos los elementos gráficos
      annotation_custom(table_plot3)
    
  # Mostrar el gráfico
    plot3
    
    ggsave("f2.png", plot3, width = 12, height = 4.15, dpi = 600)

  # Convertir el dataframe en un objeto tipo tabla usando tableGrob
    table_plot3b <- gridExtra::tableGrob(
        IGC_step2_skinnydf %>%
            dplyr::mutate(across(where(is.numeric), ~ round(., 3))) %>% 
            dplyr::select(c(item, Response, lcs, similarity_str, 
                            match_str = itemL_adj_strict_match_pos)) %>%
          dplyr::rename(response = Response) %>%
          dplyr::slice_head(n = 7) # Seleccionar las primeras 7 filas
      )

  # Crear un gráfico vacío para agregar la tabla, eliminando los márgenes
    plot3b <- ggplot() +
      theme_void() +  # Eliminar todos los elementos gráficos
      annotation_custom(table_plot3b)
    
  # Mostrar el gráfico
    plot3b
    
    ggsave("f2b.png", plot3b, width = 12, height = 4.15, dpi = 600)

```

## 2.1. Positional Analysis

Obtain the correct characters, in this case, letters, in their correct position using the `positional_accuracy()` function.


```{r, positions}

IGC_step2.1 = IGC_step2 %>% positional_accuracy(match_col = "itemL_adj_strict_match_pos",
                                          last_ID_col = "targetL")

IGC_step2.1_skinnydf = IGC_step2.1 %>% dplyr::select(-c(task_ID, correct, task_type)) %>%
            dplyr::rename(item_ID = task_item_ID, attempt = Attempt, response = Response, position = Position)

rmarkdown::paged_table(IGC_step2.1_skinnydf, options = list(rows.print = 25, align = "c"))

```

```{r include=FALSE}

# Convertir targetL a character para evitar problemas al combinar dataframes
positions_accuracy <- IGC_step2.1 %>%
  dplyr::mutate(targetL = as.character(targetL))

# Duplicar y modificar el dataframe para crear 'positions_general'
positions_general <- positions_accuracy %>%
  dplyr::mutate(targetL = "General")

# Combinar ambos dataframes
positions <- dplyr::bind_rows(positions_accuracy, positions_general)

# Especificar manualmente los niveles en el orden deseado
desired_levels <- c("3", "4", "5", "6", "7", "8", "9", "10", "11", "12",
                    "13", "14", "15", "17", "21", "22", "24", "48", "General")

# Convertir correct_pos a numérico y ordenar targetL como factor según desired_levels
positions <- positions %>%
  dplyr::mutate(correct_pos = as.numeric(correct_pos),
         targetL = factor(targetL, levels = desired_levels)) %>%
  dplyr::arrange(correct_pos, targetL)

# Definir un conjunto de linetypes que se pueda repetir
custom_linetypes <- rep(c("solid", "dashed", "dotted", "longdash", "dotdash"),
                        length.out = nlevels(positions$targetL))

# Calcular la precisión y contar el número de observaciones por grupo
plot_positions <- positions %>%
  dplyr::group_by(Position, targetL) %>%
  dplyr::summarize(acc = mean(correct_pos, na.rm = TRUE)*100,
            n = dplyr::n()) %>%
  ggplot(aes(x = as.numeric(Position), y = acc, group = targetL,
             fill = targetL, color = targetL, lty = targetL)) +
  geom_line(size = 0.70, alpha = 0.6) +
  geom_point(aes(size = n), shape = 21, color = "black", alpha = 0.6) +
  scale_linetype_manual(values = custom_linetypes) +
  theme(panel.border = element_rect(colour = "black", fill = NA),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank()) +
  ylab("Positional accuracy (%) ") +
  xlab("Position") +
    ylim(0, 100) +  # Aseguramos que el eje y muestre de 0 a 100
  guides(fill = guide_legend(title = "Word Length"),
         lty = guide_legend(title = "Word Length"),
         color = guide_legend(title = "Word Length"),
         size = guide_legend(title = "Datapoints"))  +
    theme_grey(base_size = 12) +
    theme(panel.grid = element_blank(),
          axis.line = element_line(colour = "black"),
          plot.title = element_text(hjust = 0.5),
          legend.position = "right",
          axis.title.x = element_text(margin = margin(t = 10)),
          axis.title.y = element_text(margin = margin(r = 10)))

plot_positions

ggsave("pa.png", plot = plot_positions, width = 10, height = 5)

```

This is a file (generated following the procedure described by [Dueñas Lerín] (https://duenaslerin.com/diccionario-palabras-espanol-en-texto-script/)) containing all the words in Spanish as available in the RAE dictionary. It can be downloaded from the author's page at https://github.com/JorgeDuenasLerin/diccionario-espanol-txt.

```{r load some required files: word2vec model}

m_w2v = word2vec::read.word2vec(file = file.choose(), normalize = F)

```

This is a file (generated using the word2vec algorithm by [Cardellino](https://crscardellino.github.io/)) containing the embeddings of 1.5 billion words. It can be downloaded from the [author's page]( https://github.com/JorgeDuenasLerin/diccionario-espanol-txt) or in another mirror at [Github](https://github.com/dccuchile/spanish-word-embeddings), where others corpuses can be assessed; further details  provided by the author [here](https://crscardellino.github.io/SBWCE/)


```{r, get indexes}

IGC_step2_clean = IGC %>%
                        separate_responses(
                          col_name = "final_response",
                          separate_with = ", ") %>% 
                        get_attempts(
                          first_production = Attempt_1, drop_blank_spaces = T)  %>%
                        dplyr::select(task_ID, ID, task_item_ID, task_type, item, Response, RA, Attempt) %>% 
                        get_formal_similarity(item_col = "item", response_col = "Response",
                          attempt_col = "Attempt",
                          group_cols = c("ID", "task_item_ID"))

IGC_step2clean_skinnydf = IGC_step2_clean %>% dplyr::select(-c(task_ID, task_type))

rmarkdown::paged_table(IGC_step2clean_skinnydf, options = list(rows.print = 25, align = "c"))

# remove some values leaving NAs to check that the functions work correctly

IGC_step2_cleanNA = IGC_step2_clean %>%
  dplyr::mutate(
    Response = dplyr::if_else(dplyr::row_number() == 2, NA_character_, Response),
    item = dplyr::if_else(dplyr::row_number() == 3, NA_character_, item)
  )

IGC_step3 <- IGC_step2_cleanNA %>%
  check_lexicality(item_col = "item", response_col = "Response", criterion = "database") %>%
  get_formal_similarity(item_col = "item", response_col = "Response",
                          attempt_col = "Attempt",
                          group_cols = c("ID", "task_item_ID")) %>%
  get_semantic_similarity(item_col = "item", response_col = "Response", model = m_w2v)

# Compute accessed col

IGC_step3 = IGC_step3 %>% 
    dplyr::mutate(accessed = dplyr::if_else(Response == item, 1, 0))

IGC_step3_skinnydf = IGC_step3 %>% dplyr::select(-c(task_ID, task_type))

rmarkdown::paged_table(IGC_step3_skinnydf, options = list(rows.print = 25, align = "c"))

```

Proceed with errors classification

```{r, classification}

IGC_step4 <- IGC_step3 %>% classify_errors(access_col = "accessed", 
                                           RA_col = "RA",
                                           response_col = "Response", 
                                           item_col = "item",
                                           also_classify_RAs = T) %>%
  dplyr::mutate(general_ID = dplyr::row_number())

IGC_step4_print_skinny <- IGC_step4 %>%
  dplyr::filter(general_ID %in% c(2636, 790, 2570, 2492, 866, 300)) %>%
  dplyr::select(general_ID, ID, task_item_ID, item, Response, RA, 
                Attempt, lexicality, cosine_similarity, nonword:no_response) %>% 
  dplyr::rename(w2v_cos = cosine_similarity, item_ID = task_item_ID, 
         attempt = Attempt, response = Response) %>% 
  dplyr::select(-c(general_ID, no_response))

rmarkdown::paged_table(IGC_step4_print_skinny, options = list(rows.print = 9, align = "c"))


  # Convertir el dataframe en un objeto tipo tabla usando tableGrob
      table_plot4 <- gridExtra::tableGrob(
        IGC_step4_print_skinny %>%
          dplyr::slice_head(n = 7)  # Seleccionar las primeras 7 filas
      )

    # Crear un gráfico vacío para agregar la tabla, eliminando los márgenes
    plot4 <- ggplot() +
      theme_void() +  # Eliminar todos los elementos gráficos
      annotation_custom(table_plot4)
    
    # Mostrar el gráfico
    plot4
    
    ggsave("f3.png", plot4, width = 13, height = 4.15, dpi = 600)

```
